---
title: "Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# library(CellDART)
```

# 0. Install required packages
## Install CellDART
```{r}
# if (!requireNamespace("CellDART", quietly = TRUE))
#   devtools::install_github("mexchy1000/CellDART")
```


# 1. Example of using function "pred_cellf_celldart"
```{r}
# library(SeuratObject)
# 
# # Find the directory for active script file (file_path)
# file_path <- rstudioapi::getSourceEditorContext()$path
# file_path <- strsplit(file_path, split=.Platform$file.sep)
# file_path <- paste(file_path[[1]][-length(file_path[[1]])],
#                     collapse=.Platform$file.sep)
# 
# # Set working directory
# setwd(file_path)
# 
# # Make output folder
# output_folder_name <- 'CellDART_output'
# if (!file.exists(output_folder_name)){
#   dir.create(output_folder_name)
# }
```

## Load single-cell and spatial datasets
### Load single-cell dataset (RDS file with Seurat object): GSE115746 (mouse single cell: ALS and VISp)
```{r}
# sc_data <- readRDS('sc_data.rds')
```


## Load spatial dataset (RDS file with Seurat object): 10X genomics data repository
### V1_Mouse_Brain_Sagittal_Anterior & V1_Mouse_Brain_Sagittal_Posterior
```{r}
# sp_data <- readRDS('sp_data.rds')
```

## Check the size of spatial dataset
```{r}
# dim(sp_data)
```


## Set the number of pseudospots: 5 times the number of spatial spots
```{r}
# npseudo <- 5*dim(sp_data)[2]
```

## Perform CellDART analysis
### Explanation of the function
```{r}
help(CellDART)
```


### Using conda environment (environment will be automatically installed in Linux distributions)
#### If using Windows, then install conda environment first and then run the function below with python.install = F

```{r}
# sp_data_cellf <- CellDART::pred_cellf_celldart(sp_data, sc_data, outdir = file.path(output_folder_name),
#                                                sp_subset=F, spot.cluster.name='seurat_clusters',
#                                                spot.cluster.of.interest=NULL,
#                                                env.select='conda',python.install=T,
#                                                python_path=NULL, env.name='CellDART',
#                                                gpu=TRUE, metadata_celltype='cell_subclass',
#                                                num_markers=20, seed_num=0,
#                                                nmix=8, npseudo=npseudo, alpha=0.6,alpha_lr=5,
#                                                emb_dim=64,batch_size=512,n_iterations=3000, init_train_epoch=10)
```


### Using virtual environment (environment will be automatically installed in Linux distributions)
#### Not recommended for Windows
```{r}
# sp_data_cellf <- CellDART::pred_cellf_celldart(sp_data, sc_data, outdir = file.path(output_folder_name),
#                                                sp_subset=F, spot.cluster.name='seurat_clusters',
#                                                spot.cluster.of.interest=NULL,
#                                                env.select='virtual',python.install=T,
#                                                python_path=NULL, env.name='CellDART',
#                                                gpu=TRUE, metadata_celltype='cell_subclass',
#                                                num_markers=20, seed_num=0,
#                                                nmix=8, npseudo=npseudo, alpha=0.6,alpha_lr=5,
#                                                emb_dim=64,batch_size=512,n_iterations=3000, init_train_epoch=10)
```


### Save seurat object with cell fraction
```{r}
# saveRDS(sp_data_cellf, file.path(output_folder_name, 'sp_data_cellf.rds'))
```

# 2. Visualization of spatial cell fraction
### Remove '_cellf' from the column names cell fraction metadata
```{r}
# cellf.data <- sp_data_cellf@meta.data
# cellf.data.colname <- sapply(colnames(cellf.data), function(x){
#   if (grepl('_cellf',x)){return(strsplit(x, split='_cellf')[[1]][1])}
#   else {return(x)}
# })
# sp_data_cellf.mod <- sp_data_cellf
# colnames(sp_data_cellf.mod@meta.data) <- cellf.data.colname
```

### Visualize the layer-specific excitatory neuons
```{r}
# cell_types <- c("L2.3.IT","L4","L5.IT","L5.PT","L6b","L6.CT","L6.IT")
# p <- Seurat::SpatialFeaturePlot(sp_data_cellf.mod, features = cell_types, 
#                                 ncol = 4, alpha=0.6, combine = FALSE)
# patchwork::wrap_plots(p, ncol = 8)
```
